---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
activity<- read.csv("activity.csv")
head(activity)
tail(activity)
```


## What is mean total number of steps taken per day?
```{r}
steps<-setNames(aggregate(activity$steps,by=list(activity$date),FUN=mean),c("Date","Mean Steps"))
steps[,1]<-as.numeric(steps[,1])
library(ggplot2)
plot(steps[,1],steps[,2],xlab="Day",ylab="Mean Steps",main="Mean total steps taken per day",type="h")
```


## What is the average daily activity pattern?
```{r}
daily<-setNames(aggregate(activity$steps,by=list(activity$interval),FUN=mean,na.rm=T),c("Interval","Steps"))
library(ggplot2)
plot(daily[,1],daily[,2],type="l",xlab="Interval",ylab="Steps",main="Average Daily Activity Pattern")
```


## Imputing missing values
```{r}
naValues<-which(is.na(activity$steps))
meanSteps<-mean(steps$'Mean Steps',na.rm=T)
activity2<-activity
for (n in naValues) activity2$steps[n]<-meanSteps
steps2<-setNames(aggregate(activity2$steps,by=list(activity2$date),FUN=mean),c("Date","Mean Steps"))
steps2[,1]<-as.numeric(steps2[,1])
library(ggplot2)
plot(steps2[,1],steps2[,2],xlab="Day",ylab="Mean Steps",main="Mean total steps taken per day",type="h")
```

#### Mean and Median total number of steps Excluding NA values
```{r}
mean<-sum(steps$'Mean Steps',na.rm=T)/length(steps$'Mean Steps')
print(mean)
median(steps$'Mean Steps',na.rm=T)
```

#### Mean and Median total number of steps Including NA values that are replaced
```{r}
mean(steps2$'Mean Steps')
median(steps2$'Mean Steps')
```
## Are there differences in activity patterns between weekdays and weekends?

```{r}
weekday.or.weekend <- function(date) {
    day <- weekdays(date)
    if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
        return("weekday")
    else if (day %in% c("Saturday", "Sunday"))
        return("weekend")
    else
        stop("invalid date")
}

activity2$date <-as.Date(activity2$date)
x<-character(0)
for (i in 1:length(activity2$date)){ x[i]<-weekday.or.weekend(activity2$date[i])}

activity2<-cbind(activity2,x)
weekday<-activity2[activity2$x=='weekday',]
#aggStep<-aggregate(x = activity2$steps,by = list(activity2$date),FUN = mean)

aggInterval<-aggregate(x = weekday$steps,by = list(weekday$interval),FUN = sum)
plot(x = aggInterval$Group.1,y = aggInterval$x,type = "l",xlab="Intervals",ylab="Steps",main="Weekdays")

weekend<-activity2[activity2$x=='weekend',]
aggInterval<-aggregate(x = weekend$steps,by = list(weekend$interval),FUN = sum)
plot(x = aggInterval$Group.1,y = aggInterval$x,type = "l",xlab="Intervals",ylab="Steps",main="WeekEnds")

```

